---
title: "P3 - Summary Statistics for Numerical Variables"
author: "Jose Rico"
date: "9/10/2019"
output:
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
    df_print: kable
---

# Packages Needed to be Loaded with `library` function
Note: They have already been installed in this project
```{r, message=F}
library(ggplot2)
library(dplyr)
library(readr)
library(purrr)
library(kableExtra)
library(plotly)
```

# Assignment Details
Choose one of the survey questions from the survey that you are interested in exploring the results.

Link to Survey Results - https://docs.google.com/spreadsheets/d/1cvoTnrb1FLaiHcXBEt5q8oG4ZpKp-dc6zaTs6QX253w/edit?usp=sharing

For the survey question complete the following tasks;

* Create at least 2 visualizations of the distribution
* Calculate summary statistics
* Write your own summary describing the distribution
  + Shape (modalities)
  + Skew
  + Center (mean/median)
  + Outliers
  + Middle 50%
* **Exceptional** grade is possible if you break the variable down with a categorical variable

# Load Data
In order to load data, I suggest the following procedure;  

* Copy the Google sheet, clean up response (so they are numerical) and rename the columns/variable
(so they are one word, not questions)  

* Publish to Web (under file)  

* Enter link inside the parenthesis

```{r, loading data from Google}
# survey <- read_csv("PUBLISHED GOOGLE FILE GOES HERE")  #  code ignored to knit - erase front #
```

# Question 1 - 

## Visualization

## Numerical Summary

## Narrative Summary


